Energy-based automatic detection on zebra finch songs

### 2023-05-24 manuscript/manuscript.Rmd
many changes

---
title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

@@

```{r word count, eval = TRUE, echo = FALSE, results='asis'}

@@

head(split_info)
```

@@

@@

Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection. These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by *ohun*, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources, rather than a direct comparison of performance between Raven Pro and *ohun*.

@@

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")

@@

diagnose_detection(reference = test_ref, detection = detec_test)
```

@@

sub_opt_det_train
```

>>>>

---
title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
author:
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

# options to customize chunk outputs
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors", # for code line numbers
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = FALSE,
  warning = FALSE
 )
```

@@

```{r word count, eval = FALSE, echo = FALSE, results='asis'}

@@

as.matrix(split_info[1:6,])
```

@@

dd <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

as.matrix(dd)

@@

as.matrix(diag_raven[1:10, ])

```
Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection. These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by *ohun*, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources, rather than a direct comparison of performance between Raven Pro and *ohun*.

@@

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")

@@

dd <- diagnose_detection(reference = test_ref, detection = detec_test)

as.matrix(dd)
```

@@

as.matrix(sub_opt_det_train)
```

### 2023-05-24 manuscript/manuscript_v01.Rmd
many changes

---
title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
author:
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

# options to customize chunk outputs
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors", # for code line numbers
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = FALSE,
  warning = FALSE
 )
```

@@

```{r, eval = FALSE, echo = FALSE}
source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)

@@

@@

raven_detec <- read.csv("combined_raven_detection.csv")

diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

@@

raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")

dd <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

as.matrix(dd)

@@

The output shows the indices described above. Plus three additional metrics specific to sound event detection: 'split positives', 'merge positives' and 'overlap to true positives'. 'Split positives' is the number of reference sounds overlapped by more than one detection, 'merge positives' the number of detected sounds overlapping with another detection and 'overlap to true positives' quantifies the mean overlap between detections and reference signal (1 means complete overlap). The function also allows to detail those indices by sound file. Here we show the first 10 files:

@@

```{r, eval = TRUE, echo = TRUE}

@@

as.matrix(diag_raven[1:10, ])

```
Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection. These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by *ohun*, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources, rather than a direct comparison of performance between Raven Pro and *ohun*.

@@

# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f1.score, decreasing = TRUE), ]

# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f1.score, opt_detec_train$templates, FUN = rank) <= 4, ]

@@

    X <- opt_detec_train[opt_detec_train$templates == x, ]
    X <- X[order(X$f1.score, decreasing = TRUE), ]
    return(X[1:2, ])
    })

@@

kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f1.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.", digits = 3)

@@

We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F1.SCORE).

```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values"}

agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
    geom_line() + 
    geom_point() + 
    scale_color_viridis_d(end = 1, labels = c("Centroid", "1st", "2nd", "3rd"), alpha = 0.7) + 
    labs(x= "Cross-correlation threshold", y = "F1 score", color = "Templates")

@@

In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F1.SCORE). Hence, we will use this template for detecting calls on the rest of the data. The following code extracts this template from the reference annotation table and use it to find inquiry calls on the testing data set:

@@

dd <- diagnose_detection(reference = test_ref, detection = detec_test)

as.matrix(dd)
```

@@

opt_det_train <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10)
)

@@

# subset with highest performance
opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

@@

# subset with highest performance
sub_opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

as.matrix(sub_opt_det_train)
```

@@

best_param <- opt_det_train[which.max(opt_det_train$f1.score), ]

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10)
)
```

@@

>>>>

---
title: 'ohun: an R package for diagnosing and optimizing automated sound event detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: no
    toc_depth: '3'
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

# options to customize chunk outputs
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors", # for code line numbers
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = TRUE,
  warning = FALSE
 )
```

@@

```{r, eval = FALSE, echo = FALSE}
source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)

@@

raven_detec <- raven_detec[, c("sound.files", "selec", "start", "end", "bottom.freq", "top.freq", "tuning_parameters")]

@@

```{r, eval = FALSE, echo = TRUE}
raven_detec <- read.csv("combined_raven_detection.csv")

head(raven_detec)
```

```{r, eval = TRUE, echo = FALSE}
raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")

head(raven_detec)
```

```{r, eval = FALSE, echo = TRUE}
diagnose_detection(reference = manual_ref_tae,
                   detection = raven_detec,
                   by = "tuning_parameters",
                   min.overlap = 0.5)

@@

diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", min.overlap = 0.5)

@@

The output shows the indices described above. Plus three additional metrics specific to sound event detection: 'split positives', 'merge positives' and 'overlap to true positives'. 'Split positives' is the number of reduntant detections (i.e. those overlapping sounds already overlapped by other detections) by more than one detection, 'merge positives' the number of detected sounds overlapping with another detection and 'overlap to true positives' quantifies the mean overlap between detections and reference signal (1 means complete overlap).

diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

head(diag_raven, 10)

head(diag_raven, 10)

The function also allows to detail those indices by sound file. Here we show the first 10 files:

@@

@@

head(diag_raven, 10)

Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize detection. These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by *ohun*, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources, rather than a direct comparison of performance between Raven Pro and *ohun*.

@@

# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f.score, decreasing = TRUE), ]

# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f.score, opt_detec_train$templates, FUN = rank) <= 4, ]

@@

    X <- opt_detec_train[opt_detec_train$templates == x, ]
    X <- X[order(X$f.score, decreasing = TRUE), ]
    return(X[1:2, ])
    })

@@

kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.", digits = 3)

@@

We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F.SCORE).

```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F.SCORE. Shows the changes in F1 score across the range of threshold values"}

agg_f1 <- aggregate(f.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f.score, group = templates, color = templates)) +
    geom_line() + 
    geom_point() + 
    scale_color_viridis_d(end = 1, labels = c("Centroid", "1st", "2nd", "3rd"), alpha = 0.7) + 
    labs(x= "Cross-correlation threshold", y = "F1 score", color = "Templates")

@@

In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F.SCORE). Hence, we will use this template for detecting calls on the rest of the data. The following code extracts this template from the reference annotation table and use it to find inquiry calls on the testing data set:

@@

@@

opt_det_train <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10),
  path = data_path
)

```{r, eval = FALSE, echo = FALSE}
write.csv(opt_det_train , file = "./data/processed/taeniopygia/detection_optimization.csv")

@@

# subset with highest performance
opt_det_train <- opt_det_train[order(opt_det_train$f.score, decreasing = TRUE), ]

@@

# subset with highest performance
sub_opt_det_train <- opt_det_train[order(opt_det_train$f.score, decreasing = TRUE), ]

sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

head(sub_opt_det_train, 10)
```

@@

best_param <- opt_det_train[which.max(opt_det_train$f.score), ]

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10),
  path = data_path, cores = 10 
)
```

@@

Using macro averaging:

```{r, diagnose 2, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref,
                   detection = det_test,
                   by.sound.file = FALSE,
                   macro.average = TRUE)

### 2023-05-24 scripts/study_case_spixs_disc-winged_bat_calls.Rmd
update figs and manuscript
many changes

---
title: <center><font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font></center>
subtitle: <center><font size="4"><b>ohun</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

@@

### Load packages
```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, message = FALSE, warning=FALSE}

@@

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
```

```{r, message=FALSE, warning=FALSE}
library(ohun)

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
```{r Create catalogs, eval = FALSE}

@@

# create catalogs
catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL)

@@

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}

knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

@@

### Get diverse set of templates
```{r get templates, eval = TRUE, echo = TRUE, fig.height=7}
# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.sub.spaces = 3)
```

```{r get templates 2, eval = FALSE, echo = TRUE}
# create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, confirm.extended = FALSE, path = data_path)

@@

# create catalogs
catalog(X = templates_est, flim = c(10, 45), nrow = 2, ncol = 2, ovlp = 90, height = 10, width = 15, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), img.prefix = "templates")

@@

knitr::include_graphics(normalizePath(file.path(data_path, "templates-Catalog_p1.jpeg")))

>>>>

---
title: <font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font>
subtitle: <font size="4"><b>ohun</b></font>
author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
date:  "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

<div class="alert alert-warning">

### Data set

The data used in this tutorial (annotations and recordings) can be downloaded from here:

- [https://figshare.com/ndownloader/files/38431409](https://figshare.com/ndownloader/files/38431409)

@@

### Package installation

Install/load the package from CRAN as follows:

# From CRAN would be
install.packages("ohun")

#load package
library(ohun)

To install the latest developmental version from [github](https://github.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes):

```{r, eval = FALSE}
remotes::install_github("maRce10/ohun")

#load package
library(ohun)

### Load additional packages
```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, message = FALSE, warning=FALSE}

@@

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
```

```{r, message=FALSE, warning=FALSE}

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` is a image file in the current working directory):
```{r Create catalogs, eval = FALSE}

@@

# create catalogs
catalog(X = high_snr, flim = c(10, 45), nrow = 6, ncol = 5, ovlp = 90, height = 10, width = 20, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = FALSE, pal = viridis, collevels = seq(-100, 0, 5), labels = NULL, path = data_path)

@@

![](../data/raw/thyroptera/Catalog_p1.jpeg)

@@

### Get diverse set of templates
```{r get templates, eval = TRUE, echo = TRUE, fig.height=7}

# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), fast = TRUE, ovlp = , wl = 200, n.sub.spaces = 3)

```{r get templates 2, eval = FALSE, echo = TRUE}

# create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, confirm.extended = FALSE, path = data_path)

@@

# create catalogs
catalog(X = templates_est, flim = c(10, 45), nrow = 2, ncol = 2, ovlp = 90, height = 10, width = 15, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE, spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-100, 0, 5), img.prefix = "templates", path = data_path)

@@

knitr::include_graphics("./data/raw/thyroptera/templates-Catalog_p1.jpeg")

### 2023-05-24 scripts/study_case_spixs_disc-winged_bat_calls.qmd
many changes

---
title: <font size="6"><b>Template-based automatic detection on Spix's disc-winged bat social calls</b></font>
subtitle: <font size="4"><b>ohun</b></font>
author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
date:  "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

opts_chunk$set( fig.width = 8, fig.height = 3.5)

@@

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
```

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` is a image file in the current working directory):
```{r Create catalogs, eval = FALSE}

@@

![](../data/raw/thyroptera/Catalog_p1.jpeg)

@@

```{r catalog templates, out.width = "70%", echo = FALSE, fig.align= "center"}

knitr::include_graphics("./data/raw/thyroptera/templates-Catalog_p1.jpeg")

@@

corr_templ_train <- template_correlator(
  templates = templates_est,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  parallel = paral, 
  hop.size = 10, 
  ovlp = 70
  )

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.1, 0.9, 0.05), 
  parallel = paral
  )

@@

### Create plot with threshold vs f1 score
```{r, eval = TRUE, echo = FALSE}

@@

agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()

@@

### Check subset with highest performance (sorted by f1 score)

@@

opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

@@

# subset with highest performance
opt_detec_train <- opt_detec_train[order(opt_detec_train$f1.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

@@

dd$f1.score

@@

- Good performance on detecting Spix's disc-winged social calls: F1 score was `r round(dd$f1.score, 3)` for the training data set and `r round(max(opt_detec_train$f1.score), 3)` for the testing data

>>>>

---
title: Template-based automatic detection on Spix's disc-winged bat social calls
subtitle: ohun
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
date: "`r Sys.Date()`"
toc: true
toc-depth: 2
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: show
    code-tools: true
    css: qmd.css
---

@@

opts_chunk$set( fig.width = 8, fig.height = 3.5, warning = FALSE)

@@

source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/summarize_diagnostic.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
```

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set (note that the output of the function `catalog` are image files saved in the sound file directory):
```{r Create catalogs, eval = FALSE}

@@

![](./img/Catalog_thyroptera.jpeg)

@@

![](./img/templates-Catalog_p1.jpeg)

@@

corr_templ_train <- template_correlator(
  templates = templates_est,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  cores = paral, 
  hop.size = 10, 
  ovlp = 70
  )

opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.1, 0.9, 0.05), 
  cores = paral
  )

@@

### Create plot with threshold vs F score
```{r, eval = TRUE, echo = FALSE}

@@

agg_f <- aggregate(f.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f, aes(x = threshold, y = f.score, group = templates, color = templates)) +
geom_line() + geom_point() + scale_color_viridis_d(end = 1) + theme_classic()

@@

### Check subset with highest performance (sorted by F score)

@@

opt_detec_train <- opt_detec_train[order(opt_detec_train$f.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

@@

# subset with highest performance
opt_detec_train <- opt_detec_train[order(opt_detec_train$f.score, decreasing = TRUE), ]

sub_opt_detec_train <- opt_detec_train[1:10, c("templates","threshold", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

@@

dd$f.score

@@

- Good performance on detecting Spix's disc-winged social calls: F score was 0.953 for the training data set and 0.955 for the testing data

### 2023-05-24 scripts/study_case_zebra_finch_songs.Rmd
update figs and manuscript
many changes

---
title: <center><font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font></center>
subtitle: <center><font size="4"><b>ohun</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

@@

### Load package
```{r libraries, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

@@

knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

@@

>>>>

---
title: <font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font>
subtitle: <font size="4"><b>ohun</b></font>
author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

<div class="alert alert-warning">

### Data set

The data used in this tutorial (annotations and recordings) can be downloaded from here:

- [https://figshare.com/ndownloader/files/38431340](https://figshare.com/ndownloader/files/38431340)

@@

### Package installation

Install/load the package from CRAN as follows:

```{r, eval = FALSE}

# From CRAN would be
install.packages("ohun")

#load package
library(ohun)

To install the latest developmental version from [github](https://github.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes):

```{r, eval = FALSE}
remotes::install_github("maRce10/ohun")

#load package
library(ohun)

### Load additional packages
```{r libraries, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

@@

# knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

@@

![](../data/raw/taeniopygia/Catalog_p1.jpeg)
&nbsp;

### 2023-05-24 scripts/study_case_zebra_finch_songs.qmd
many changes

---
title: <font size="6"><b>Energy-based automatic detection on zebra finch songs</b></font>
subtitle: <font size="4"><b>ohun</b></font>
author: <font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font>
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

@@

knitr::opts_knit$set(root.dir = normalizePath(".."))

knitr::opts_chunk$set(include = TRUE, eval = TRUE)

@@

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set
```{r Create catalogs, eval = FALSE}

@@

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "center"}

# knitr::include_graphics(normalizePath(file.path(data_path, "Catalog_p1.jpeg")))

![](../data/raw/taeniopygia/Catalog_p1.jpeg)
&nbsp;

@@

### Check subset with highest performance (sorted by f1 score)
```{r, eval = FALSE}

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

@@

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f1.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

@@

### Select the highest performance parameters based on f1 score
```{r, select best params, eval = TRUE, echo = TRUE}

best_param <- opt_det[which.max(opt_det$f1.score), ]

@@

- Good performance on detecting zebra finch songs: F1 score was 0.95 for the training data set and 0.96 for the testing data

>>>>

---
title: Energy-based automatic detection on zebra finch songs
subtitle: ohun
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
date: "`r Sys.Date()`"
toc: true
toc-depth: 2
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: show
    code-tools: true
    css: qmd.css
---

@@

knitr::opts_knit$set(root.dir = "..")

knitr::opts_chunk$set(include = TRUE, eval = TRUE, warning = FALSE)

@@

source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/summarize_diagnostic.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
```

@@

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set  (note that the output of the function `catalog` are image files saved in the sound file directory):
```{r Create catalogs, eval = FALSE}

@@

![](img/Catalog_taenopygia.jpeg)

@@

### Check subset with highest performance (sorted by F score)
```{r, eval = FALSE}

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

@@

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

@@

### Select the highest performance parameters based on F score
```{r, select best params, eval = TRUE, echo = TRUE}

best_param <- opt_det[which.max(opt_det$f.score), ]

@@

- Good performance on detecting zebra finch songs: F score was 0.95 for the training data set and 0.96 for the testing data
